\documentclass[12pt]{article}

\newif\ifsol
\soltrue

% Unicode compatibility
\usepackage{iftex}
\ifPDFTeX
  \usepackage[utf8]{inputenc}
  \usepackage[noTeX]{mmap}
  \usepackage[T1]{fontenc}
\fi
% XeTeX does not support mmap
\ifLuaTeX
  \usepackage{luatex85}
  \usepackage[noTeX]{mmap}
\fi

\setlength{\textheight}{9in}
\setlength{\textwidth}{7.1in}
\setlength{\evensidemargin}{-0.2in}
\setlength{\oddsidemargin}{-0.2in}
\setlength{\headsep}{30pt}
\setlength{\topmargin}{-0.3in}

\usepackage{amsthm,amsfonts,amsmath,xspace,tikz,varwidth,cancel,soul}
\usepackage{hyperref,verbatim}
\newtheorem{theorem}{Theorem}
\theoremstyle{definition}
\newtheorem{definition}{Definition}

\newcommand{\addmedskip}{\addvspace{\medskipamount}}

\newcommand{\addbigskip}{\addvspace{\bigskipamount}}

\pagestyle{myheadings}
\markboth{BU CAS CS 538.}{BU CAS CS 538.} 
\newcounter{problemnum}
\setcounter{problemnum}{0}
\newenvironment{problem}
     {\addbigskip \setcounter{partnum}{0}
      \noindent\stepcounter{problemnum}\textbf{Problem
                                             \arabic{problemnum}.\ }}
     {\par\addbigskip}
     
\ifsol
\newenvironment{solution}
     {\addbigskip
      \noindent\color{blue}\textbf{Solution.\ }}
     {\par\addbigskip}
\else
\newenvironment{solution}
     {\comment}
     {\endcomment}
\fi

\newcounter{partnum}
\setcounter{partnum}{0}
\newenvironment{ppart}
     {\addmedskip
      \noindent\stepcounter{partnum}\textbf{(\alph{partnum})}\ }
     {\par\addbigskip}

\newcommand{\Enc}{\mathrm{Enc}}
\newcommand{\Dec}{\mathrm{Dec}}
\newcommand{\pk}{\mathit{pk}}
\newcommand{\sk}{\mathit{sk}}
\newcommand{\qrn}{\mathit{QR}_n}
\newcommand{\qrp}{\mathit{QR}_p}
\newcommand{\qrpone}{\mathit{QR}_{p_1}}
\newcommand{\qrptwo}{\mathit{QR}_{p_2}}
\newcommand{\prg}{\mathit{J}}

\newcommand{\crt}{\mathrm{crt}}


\newcommand{\eqdef}{\buildrel{\scriptstyle{\mathit{def}}}\over{=}}
\newcommand{\rfrom}{\buildrel{\scriptstyle{\mathrm{R}}}\over{\leftarrow}}

\newcommand{\mbmod}{\,\%\,}

\newcommand{\A}{\mathcal{A}}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\G}{\mathbb{G}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\Y}{\mathcal{Y}}
\newcommand{\K}{\mathcal{K}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\R}{\mathcal{R}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\randk}{\mathsf k}
\newcommand{\randm}{\mathsf m}
\newcommand{\randc}{\mathsf c}
\newcommand{\Adv}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\SSadv}{\textrm{SS}\textsf{adv}}
\newcommand{\PRGadv}{\textrm{PRG}\textsf{adv}}
\newcommand{\threebitadv}{\textrm{3Bit}\textsf{adv}}
\newcommand{\PRFadv}{\textrm{PRF}\textsf{adv}}
\newcommand{\MACadv}{\textrm{MAC}\textsf{adv}}
\newcommand{\CPAadv}{\textrm{CPA}\textsf{adv}}
\newcommand{\BCadv}{\textrm{BC}\textsf{adv}}
\newcommand{\CRadv}{\textrm{CR}\textsf{adv}}
\newcommand{\DDHprimeadv}{\textrm{DDH}'\textsf{adv}}
\newcommand{\DDHadv}{\textrm{DDH}\textsf{adv}}
\newcommand{\DDHprime}{DDH$'$}

\newtheorem{lemma}{Lemma}

\newcommand{\indist}{  \mathrel{\vcenter{\offinterlineskip
  \hbox{$\sim$}\vskip-.35ex\hbox{$\sim$}\vskip-.35ex\hbox{$\sim$}}}}

\newcommand{\samples}{\xleftarrow{R}}

\newcommand{\hdl}{H_{\mathrm{dl}}}

%%% BEGIN MIKE ROSULEK'S MACROS
\newcommand{\codebox}[1]{%
        \begin{varwidth}{\linewidth}%
        \begin{tabbing}%
            ~~~\=\quad\=\quad\=\quad\=\kill
            #1
        \end{tabbing}%
        \end{varwidth}%
}

\newcommand{\fcodebox}[1]{%
    \fcolorbox{black}{white}{\codebox{#1}}%
}

\newcommand{\titlecodebox}[2]{%
    \fboxsep=0pt%
    \fcolorbox{black}{black!10}{%
        \begin{varwidth}{\linewidth}%
        \centering%
        \fboxsep=3pt%
        \colorbox{black!10}{#1} \\
        \colorbox{white}{\codebox{#2}}%
        \end{varwidth}%
    }
}



\newcommand{\link}{\diamond}
\newcommand{\lib}[1]{\ensuremath{\L_{\textsf{#1}}}\xspace}
\newcommand{\subname}[1]{\ensuremath{\textsc{#1}}\xspace}

%%% END OF MIKE ROSULEK'S MACROS


\begin{document}
\begin{center}
\Large{\textbf{CAS CS 538.  \ifsol Solutions to \fi Problem Set 8}}\\
\smallskip
\large{\textbf{Due electronically via Gradescope, Monday April 6, 2026 11:59pm}}\\
\large{\textbf{\textit{Om Khadka, U51801771}}}
\end{center}

\noindent
\textbf{How to submit:}  See instructions on Problem Set 1. 
 


 \paragraph{Trapdoor Hash Functions.}


 \noindent
For this problem, we will augment collision-resistant hashing with a useful property.
 
We will consider keyed collision-resistant hash functions that take a hashing key $\pk$ and two input values (which we will call $(a, b)$), and output one value. We will now augment these hash functions with an additional property: a trapdoor secret key $\sk$ that enables easy collision finding. 

We formalize this property as follows. We add a key generation algorithm
 $G$ that outputs a public hashing key $\pk$ and a trapdoor key $\sk$. There is also an efficient algorithm
 $T$ to find collisions given $\sk$: specifically,  $T(\sk, a_1, b_1, a_2)$, outputs $b_2$ such
 that $H(\pk, a_1, b_1)=H(\pk, a_2, b_2)$.  We still require that the hash function remains collision resistant for anyone who is given $\pk$ but not $\sk$.


 In the next two problems you will build trapdoor hash functions. Of course, if a hash function has a trapdoor, then the person who generated the key will be able to find collisions.
 
 \begin{problem} (30 points) \label{dltrapdoor}
Recall the DL-based family in Section 8.5.1 of the textbook. Let $q$ be a Sophie Germain prime, so that $p=2q+1$ is also prime. Let  $x, y\in\qrp$ and $a, b\in \{1, 2, \dots, q\}$. The public key is $(x, y)$ and $H((x, y), a, b)=\mathrm{abs} (x^a y^b\bmod p)$, where $\mathrm{abs}(z)$ returns $z$ if $z<p/2$ and $-z$ otherwise.

Show that $H$ is actually a trapdoor hash
 family.  In other words, demonstrate an algorithm $G$ creates $\pk=(x, y)$ and some $\sk$ (your job is to figure out what $\sk$ should be), and algorithm $T$.
 \end{problem}

\begin{solution}
Given the abstract definition of the trapdoor hash func before Q1, this question is now basically making me show an actual example of how a such a func would work.
The main way I will solve this is (basically) through modular arithmetic.

\subsection*{$\G$ and $\T$}
Let $\G$ be the key gen for the trapdoor func, $\T$.\\

\noindent $\G()$:
\begin{itemize}
    \item Select rnd $x \in \qrp \backslash \{1\}$.
    \item Select rnd $s \in \Z_p^*$.
    \item Compute $y$ as: $y = x^s \mod p$.
    \item $\pk = (x, y), \sk = s$.
\end{itemize}

\noindent $\T(\sk, a_1, b_1, a_2)$:
\begin{itemize}
    \item Parse $\sk$ as $s$.
    \item Computes $b_2$ as: $b_2 = b_1 + s^{-1}(a_1 - a_2)$.
    \item Return $b_2$.
\end{itemize}

\subsection*{Correctness}
Recall that $H((x, y), a, b) = abs(x^a y^b \mod p)$. There is an important property of this $abs$ func that I need to say first-off:
\begin{itemize}
    \item Because $x, y \in \qrp$, and $\qrp$ is multiplpicatively closed, then $\forall a, b, x^a y^b \in \qrp$.
    \item $q$ is odd. That means $p \equiv_4 3$. Note that $-1$ can only be a quad. residue $\mod p \iff p \equiv_4 1$. Therefore, $-1 \notin \qrp$.
    \item Therfore, if $z_1, z_2 \in \qrp, abs(z_1) = abs(z_2) \iff z_1 \equiv_p z_2$. Basically, the case that $z_1 \equiv -z_2$ is impossible.
    \begin{itemize}
        \item Therefore we can just reduce the equation we need to solve from $abs(\cdot) \equiv_p abs(\cdot)$ to just $x^{a_1}y^{b_1} \equiv_p x^{a_2}y^{b_2}$.
    \end{itemize}
\end{itemize}

\noindent Firstly we can just sub $y$ for $x^s \mod p$:
\begin{equation*}
    x^{a_1 + sb_1} \equiv_p x^{a_2 + sb_2}
\end{equation*}

\noindent Now, from $\T$, we get $b_2 \equiv_q b_1 + s^{-1}(a_1 - a_2)$. By just multiplying both sides w/ $s$:
\begin{equation*}
    sb_2 \equiv_q sb_1 + a_1 - a_2 \to a_1 + sb_1 \equiv_q a_2 + sb_2
\end{equation*}

\noindent Now, note that $x, q$ are both prime, and that $x$ has order $q$. That means that exponents of $x$ are equivalent $\mod q$ iff the powers are equivalent modulo $p$:
\begin{equation*}
    x^{a_1 + sb_1} \equiv_p x^{a_2 + sb_2} \to a_1 + sb_1 \equiv_q a_2 + sb_2
\end{equation*}

\noindent This equation holds perfectly. It is also the exact same output as from $\T$, which is assumed to output the collision of $b_2$. Therfore, this shows how $H(\pk, a_1, b_1) = H(\pk, a_2, b_2)$ if you run $\T(\sk, a_1, b_1, a_2)$.

\subsection*{Game}
Now for the proof that $\G, \T$ holds, I will make an adversary game, with a MITM. The MITM, $\B$, will try to simluate a discrete log (DL) game for $\A$, which is playing against the CRHF.\\

\noindent Let $\A$ be an eff. adversary, that can find collisions within $H$, w/ advantage $\epsilon$, when only given $\pk$. $\B$ will then also be another efficent adv., solving the DL problem in $\qrp$ with the same probability.

\paragraph{DL Challenge} $\B$ firstly gets $(p, g, \Y)$, where $g$ generates $\qrp$ and $\Y \in \qrp$. The goal is to find $s \text{ such that } g^s \equiv_p \Y$.

\paragraph{Execution} $\B$ will set $\pk \rightarrow (g, \Y)$. This distribution is the same as a valid public key, outputted by $\G$. Then, $\B$ will give $\pk$ to $\A$. Finally, $\A$ will output $(a_1, b_1), (a_2, b_2)$, such that they cause a collision when both plugged into $H$.

\paragraph{Extracting the Key} By the properties of $abs$ from before, we are just solving for:
\begin{equation*}
    g^{a_1}\Y^{b_1} \equiv_p g^{a_2}\Y^{b_2}
\end{equation*}

\noindent Rearrage the terms:
\begin{equation*}
    g^{a_1 - a_2} \equiv_p \Y^{b_2 - b_1}
\end{equation*}

\noindent Let $\triangle a = (a_1 - a_2)$, and $\triangle b = (b_2 - b_1)$. Then, we're just solving for:
\begin{equation*}
    g^{\triangle a} \equiv_p \Y^{\triangle b}
\end{equation*}

\noindent Since $(a_1, b_1) \neq (a_2, b_2)$, there no way that $\triangle a \equiv_q \triangle b \equiv_q 0$. Imagine if $\triangle b \equiv 0$. Then $g^{\triangle a} \equiv_p 1$. And since $g$ has order $q$, this forces $\triangle a \equiv_p 0$ as well. This is impossible. So let's just say that $\triangle b \not\equiv_q 0$.\\

\noindent Now, since $q$ is prime, then $\triangle b$ is invertible $\mod q$.Then, $\B$ computes $(\triangle b)^{-1} \mod q$ and gets:
\begin{equation*}
    s \rightarrow \triangle a \cdot (\triangle b)^{-1} \mod q
\end{equation*}

\noindent This satisfies $\Y \equiv_p g^s$, therefore this is a DL challenge. $\B$ only wins when $\A$ finds a collision. Since DL is hard, $\A$'s success is neglgible.\\

\noindent This proves the security of the construction of $\G, \T$. $\qed$
\end{solution}


 \begin{problem}  (40 points at 10 each)
  In this problem, we will build a collision-resistant hash function based on the hardness of factoring rather on the discrete logarithm problem. 
 
 Let $q_1$ and $q_2$ be two distinct Sophie Germain primes, so that $p_1=2q_1+1$ and $p_2=2q_2+1$ are primes. Let $n=p_1 p_2$. Denote by $\qrpone$, $\qrptwo$, and $\qrn$ the groups of squares modulo $p_1$, $p_2$, and $n$, respectively. We know that $\qrpone$ and $\qrptwo$ are cyclic; thus, let $g_1\in \qrpone$ generate $\qrpone$ and $g_2\in\qrpone$ generate $\qrptwo$.

 \begin{ppart}
  Show by CRT (see notes from Discussion 8) that $\qrn$ is cyclic (the easiest way to do this is to try to construct a generator of $\qrn$ and show that it is indeed a generator by showing that its order is $q_1q_2$).  
 \end{ppart}

\begin{solution}
This solution really involves a lot with the CRT, alongside the properties of gcd/lcm's (in relation to quadratic residues).

\subsection*{Proof}
By the Chinese Remainder Theorem (CRT), since $n = p_1 p_2$ and $\gcd(p_1, p_2) = 1$, we can actually form a ring isomorphism of the following: $\mathbb{Z}_n \equiv \mathbb{Z}_{p_1} \times \mathbb{Z}_{p_2}$. Because of this, we implicity get the quadratic residue subgroups:
\[ \qrn \equiv \qrpone \times \qrptwo \]
The size of $\qrpone$ is $q_1$ and similarily for $\qrptwo$, it is $q_2$. And since $q_1, q_2$ are distinct primes, $\gcd(q_1, q_2) = 1$.\\

\noindent Also note, via the given $g_i$'s I'm given, and by using CRT, it is possible to create a \textbf{distinct} element, $g \in \mathbb{Z}_n^*$, that satisfies the following:
\begin{align*}
g &\equiv g_1 \pmod{p_1} \\
g &\equiv g_2 \pmod{p_2}
\end{align*}
Since $g_1, g_2$ are quadratic residues mod their respective primes, their CRT combination, $g$, should then also be a quadratic residue $\mod n$. Thus, $g \in \qrn$.\\

\noindent Via the laws of direct product groups, the order of an element is the LCM of the orders of its components:
\[ \mathrm{ord}_n(g) = \mathrm{lcm}\left(\mathrm{ord}_{p_1}(g_1), \mathrm{ord}_{p_2}(g_2)\right) \]
And because $g_1, g_2$ are generators, then $\mathrm{ord}_{p_1}(g_1) = |\qrpone| = q_1$ and $\mathrm{ord}_{p_2}(g_2) = |\qrptwo| = q_2$. I can sub these in to get:
\[ \mathrm{ord}_n(g) = \mathrm{lcm}(q_1, q_2) \]
And since $\gcd(q_1, q_2) = 1$, it is then the case that $\mathrm{lcm}(q_1, q_2) = q_1 q_2$. Therefore, $\mathrm{ord}_n(g) = q_1 q_2$.

\subsection*{End}
Because $\qrn$ has the cardinality of $|\qrpone| \cdot |\qrptwo| = q_1 q_2$, this proves how I can make an element, $g \in \qrn$, whose order = the group size. Note that this element is a generator. Thus, $\qrn$ is cyclic. $\qed$
\end{solution}
 

 \begin{ppart} 
 Show how to efficiently factor $n$ given $q_1 q_2$. In fact, you can factor $n$ given
 any multiple of $q_1 q_2$, but we won't show it here. Hint: treat $q_1$ and $q_2$ as unknowns. Construct a system of equations for these unknowns and solve it.


\begin{solution}
The solution to this involves the vars I'm given: the pub. modulus $n$, and the value $Q = q_1 q_2$. From this, I can factor $n$ out to its prime components, $p_1, p_2$.

\subsection*{Deriving the System of Equations}
Firstly, expand the definition of $n$ using the given relationship, $p_i = 2q_i + 1$:
\begin{align*}
n &= p_1 p_2 \\
  &= (2q_1 + 1)(2q_2 + 1) \\
  &= 4q_1 q_2 + 2q_1 + 2q_2 + 1 \\
  &= 4(q_1 q_2) + 2(q_1 + q_2) + 1
\end{align*}
Now, let $S = q_1 + q_2$ be the sum of the unknown primes. I can sub in the known value, $Q = q_1 q_2$, into the equation, getting:
\begin{equation*}
n = 4Q + 2S + 1
\end{equation*}
Now solving for $S$:
\begin{equation*}
2S = n - 1 - 4Q \implies S = \frac{n - 1 - 4Q}{2}
\end{equation*}

\subsection*{Solving the Unknowns}
This gives us a system of two equations for the two unknowns so far, $q_1, q_2$:
\begin{enumerate}
    \item $q_1 q_2 = Q$ (Product)
    \item $q_1 + q_2 = S$ (Sum)
\end{enumerate}
Note that, if two numbers have a sum $S$ and product $Q$, then they are simply the roots of the quadratic equation:
\begin{equation*}
x^2 - Sx + Q = 0
\end{equation*}
Now it is possible to solve for $x$, using the quadratic formula:
\begin{equation*}
x = \frac{S \pm \sqrt{S^2 - 4Q}}{2}
\end{equation*}
And since $q_1, q_2$ are distinct primes, then their discriminant, $S^2 - 4Q$, should then also be positive. By computing these roots, you get $\{q_1, q_2\}$.

\subsection*{End}
Once you've computed $q_1, q_2$ , you can compute the factors of $n$ by just subbing in $q_1, q_2$ into the OG equation:
\begin{equation*}
p_1 = 2q_1 + 1 \quad \text{and} \quad p_2 = 2q_2 + 1
\end{equation*}
Since all I've done is only basic arithmetic and int square roots (which is efficient, by polynomial in $\log n$), this proves how one can efficiently factor $n$. $\qed$
\end{solution}
\end{ppart}


 \begin{ppart} 
Let $k$ be the bit length of $n$. Let $g$ be a generator of $\qrn$. The public key for our hash function is $n$ and $g$. The inputs will be two values in the set $\{1, 2, \dots, q_1 q_2\}$. Let $H(a,b)=g^{a2^k+b}\bmod n$ (note that $a2^k+b$ is
 just concatenation of $a$ and $b$ as bit strings).  Show that this is a
 collision-resistant hash family under the assumption that such $n$ are hard
 to factor. Specifically, show how, given $(a_1, b_1)$ and $(a_2, b_2)$ that
 collide, one can factor $n$.  You may use without proof the fact that $n$
 can be efficiently factored given any non-zero multiple of $q_1q_2$.
 \end{ppart}

\begin{solution}
I can prove the OG claim with a adv. game. Let $\A$ be some eff. adv., that outputs a collision $(a_1, b_1) \neq (a_2, b_2)$ with some non-negligible probability. I can then use this collision to factor $n$.

\subsection*{Getting $E$}
By definition of a collision:
\begin{equation*}
H(a_1, b_1) = H(a_2, b_2) \implies g^{a_1 2^k + b_1} \equiv g^{a_2 2^k + b_2} \pmod n
\end{equation*}

\noindent Now 2(a) shows us how $g$ generates $\qrn$, and therefore how $ord_n(g) = |\qrn| = q_1 q_2$. Now note, in a cyclic group, $g^X \equiv g^Y \pmod n \iff X \equiv Y \pmod{ord_n (g)}$. Thus:
\begin{equation*}
a_1 2^k + b_1 \equiv a_2 2^k + b_2 \pmod{q_1 q_2}
\end{equation*}
Now let $\Delta a = a_1 - a_2$ and $\Delta b = b_1 - b_2$. By rearranging the yields now:
\begin{equation*}
\Delta a \cdot 2^k + \Delta b \equiv 0 \pmod{q_1 q_2}
\end{equation*}
This now implies that such an int, $E = \Delta a \cdot 2^k + \Delta b$ is a multiple of $q_1 q_2$.

\subsection*{$E$ is not 0}
Now note, the inputs satisfy $a, b \in \{1, \dots, q_1 q_2\}$, so then $|\Delta a| < q_1 q_2$ and also $|\Delta b| < q_1 q_2$. Since $k$ is the bit length of $n$, we then get $2^k > n > q_1 q_2$.\\
\indent If $E = 0$, then $|\Delta a| \cdot 2^k = |\Delta b|$. If $\Delta a \neq 0$, the left side should be $\ge 2^k > q_1 q_2$, while the right side should be $< q_1 q_2$. This is a contradiction. Thefore $\Delta a = 0$, which then ALSO forces $\Delta b = 0$. This then form another contradiction, as $(a_1, b_1) \neq (a_2, b_2)$.\\
\indent So then therfore, it must be the case that $E$ is a \textit{non-zero} int multiple of $q_1 q_2$.

\subsection*{Factoring}
Not, we're given that $n$ can be eff. factored, when provided with any non-0 multiple of $q_1, q_2$. Now, given $\A$'s collision, it is possible to compute $E = (a_1 - a_2)2^k + (b_1 - b_2)$, and then pass it to the factoring algorithm. This will then get the prime factors $p_1, p_2$ of $n$ with the SAME PROBABILITY that $\A$ finds a collision.

\subsection*{End}
So then, if finding collisions is non-neglibible, then it must be the case that finding $n$ should also be non-neglibible. By going back to the assumption that factoring products of distinct safe primes is hard, then the success probability of $\A$ must therefore be negligible.\\
\indent This proves that $H$ is a collision-resistant hash family. $\qed$
\end{solution}



 \begin{ppart} 
     Show also that this is a trapdoor hash family (like in Problem 1, you need to find what
 trapdoor information $G$ should output and what the algorithm $T$
 will do).
 \end{ppart}

 \begin{solution}
The whole premesis of this solution kinda builds off of what I did in 2(c). Basically, an adv. can only break the function \textit{only if they know $q_1 q_2$}. This is equivalent to have that secret key $\sk$. I'll just prove how obtaining $\sk$ is just as hard as normally factoring $n$.\\

\noindent I'll use the construction from Problem 2(c), but now with a key gen. algo., $G$, and a collision-finding algo, $T$, that uses a trapdoor.

\subsection*{Functions}
\noindent $\G()$
\begin{itemize}
    \item Gets rnd and unique Sophie Germain primes $q_1, q_2$, and then computes safe primes $p_1 = 2q_1+1$, $p_2 = 2q_2+1$.
    \item Computes $n = p_1 p_2$ and the group order $Q = q_1 q_2 = \frac{(p_1-1)(p_2-1)}{4}$.
    \item Chooses a generator $g$ of $\qrn$ (via CRT, as in 2(a)).
    \item Now, let $k$ be the bit length of $n$.
    \item Then, output $\pk \leftarrow (n, g)$ and $\sk \leftarrow Q$.
\end{itemize}

\noindent $\T(\sk, a_1, b_1, a_2)$:
\begin{itemize}
    \item Parses $\sk$ as $Q$.
    \item Calculates $b_2 \leftarrow \left(b_1 + (a_1 - a_2) \cdot 2^k\right) \bmod Q$.
    \item Returns $b_2$.
\end{itemize}

\subsection*{Correctness}
This is checking that $T$ outputs $b_2$, such that $H(a_1, b_1) = H(a_2, b_2)$.
Now note, $H(a,b) = g^{a 2^k + b} \bmod n$. A collision requires:
\begin{equation*}
g^{a_1 2^k + b_1} \equiv g^{a_2 2^k + b_2} \pmod n
\end{equation*}
Note that $g$ generates $\qrn$, $ord_n(g) = Q$. From cyclic group properties, $g^X \equiv g^Y \pmod n \iff X \equiv Y \pmod Q$. Therefroe, we need:
\begin{equation*}
a_1 2^k + b_1 \equiv a_2 2^k + b_2 \pmod Q
\end{equation*}
Solving for $b_2$:
\begin{equation*}
b_2 \equiv b_1 + (a_1 - a_2) 2^k \pmod Q
\end{equation*}
Note that this algo, $T$, computes exactly this value. So then, just subbing this $b_2$ back into the exponent congruence satisfies the condition, assuring that $H(a_1, b_1) = H(a_2, b_2)$.

\subsection*{End}
The hash family remains collision-resistant for anyone, given ONLY $\pk = (n, g)$. As shown in 2(c), finding a collision without knowing $Q$ will allow one to eff. factor $n$. Under the assumption that factoring is hard, an adv.'s success probability is negligible. However, the key generator holds $\sk = Q$, allowing them to bypass the hardness assumption, and to just compute colliding inputs $b_2$ via simple modular arithmetic in polynomial time.\\
\indent Therefore, this proves that $H$ is a valid trapdoor hash family. $\qed$
\end{solution}

\end{problem}


\begin{problem} (30 points)
	Show how to compute the root of an $n$-leaf Merkle tree in $O(\log n)$ space.
	More precisely, assume $n$ is a power of 2.  Given data items $x_1, x_2, \dots, x_n$ as values to be placed in the
	leaves of the tree, and a hash function $H$, write pseudocode to compute the
	root $r$ of the tree while never storing more than $2+\log n$
	hash values. Analyze the storage space used by your pseudocode to prove that is never more than $2+\log n$. Note that we do not count the storage of the data items. 
\end{problem}
  
\begin{solution}
Your solution goes here
\end{solution}
\end{document}
